---
title: "Jobs vs housing"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{jobs-vs-housing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 12, 
  fig.height = 10,
  cache = TRUE,
  comment = "#>"
)
```

```{r loadPackages}
library(viridis)
library(leaflet)
library(scales)
if (!AND(requireNamespace("grattanCharts", quietly = TRUE),
         requireNamespace("Census2016.DataPack", quietly = TRUE))) {
  tempLib <- tempdir()
  devtools::install_github("hughparsonage/grattanCharts", lib.loc = tempLib)
  install.packages("Census2016.DataPack",
                   repos = "https://hughparsonage.github.io/drat",
                   type = "source",
                   lib.loc = tempLib)
  library(grattanCharts, lib.loc = tempLib)
  library(Census2016.DataPack, lib.loc = tempLib)
} else {
  library(grattanCharts)
  library(Census2016.DataPack)
}
library(data.table)
library(hutils)
library(magrittr)
library(ASGS)
library(sp)
library(rgeos)
library(ggplot2)
library(journey.to.work1116)
```

```{r temp}
SA2_DZN_2011 <- fread("~/ABS-Data/2011-Home_SA2-Work_DZN-Count.csv")

DZN_2011_persons <-
  SA2_DZN_2011[, .(persons = sum(Count)), keyby = .(DZN_CODE11 = Work_DZN)]

DZN_2011_data <-
  as.data.table(DZN_2011@data) %>%
  .[, id := .I]

DZN_2011_centroids <- gCentroid(DZN_2011, byid = TRUE)
DZN_2011_centroids_tidy <-
  as.data.frame(DZN_2011_centroids) %>%
  as.data.table %>%
  .[, id := .I] %>%
  .[DZN_2011_data, on = "id"] %>%
  setnames(c("x", "y"), c("lon", "lat")) %>%
  .[]

DZN_2016_persons <-
  fread("../data-raw/2016-DZN-persons.csv", skip = 10) %>%
  setnames(1:2, c("DZN_CODE16", "persons")) %>%
  drop_empty_cols %>%
  .[DZN_CODE16 %enotin% c("Total",
                          "POW not applicable",
                          "POW No Fixed Address (OT)",
                          "Migratory - Offshore - Shipping (OT)")]

DZN_2016_data <-
  as.data.table(DZN_2016@data) %>%
  .[, id := .I]

DZN_2016_centroids <- gCentroid(DZN_2016, byid = TRUE)
DZN_2016_centroids_tidy <-
  as.data.frame(DZN_2016_centroids) %>%
  as.data.table %>%
  .[, id := .I] %>%
  .[DZN_2016_data, on = "id"] %>%
  setnames(c("x", "y"), c("lon", "lat")) %>%
  .[]

library(PSMA)
SYD <- geocode(number_first = 1, street_name = "Martin", street_type = "Place", postcode = 2000)
MEL <- geocode(number_first = 350, street_name = "Bourke", street_type = "Street", postcode = 3000)
BNE <- geocode(number_first = 261, street_name = "Queen", street_type = "Street", postcode = 4000)
ADL <- geocode(number_first = 141, street_name = "King William", street_type = "Street", postcode = 5000)
PER <- geocode(number_first = 378, street_name = "Wellington", street_type = "Street", postcode = 6000)

lat_lon_by_City <-
  data.table(City = c("SYD", "MEL", "BNE", "ADL", "PER"),
             GPO_lat = c(SYD$LATITUDE, MEL$LATITUDE, BNE$LATITUDE, ADL$LATITUDE, PER$LATITUDE),
             GPO_lon = c(SYD$LONGITUDE, MEL$LONGITUDE, BNE$LONGITUDE, ADL$LONGITUDE, PER$LONGITUDE))

id_by_city <-
  data.table::CJ(id = DZN_2011_centroids_tidy$id,
                 City = c("SYD", "MEL", "BNE", "ADL", "PER")) %>%
  .[lat_lon_by_City, on = "City"]

distance2GPO_by_DZN_CODE11 <-
  DZN_2011_centroids_tidy %>%
  .[id_by_city, on = "id"] %>%
  .[, distance2GPO := haversine_distance(lat, lon, GPO_lat, GPO_lon)] %>%
  .[, .(distance2GPO = min(distance2GPO),
        City = City[which.min(distance2GPO)]), keyby = c("DZN_CODE11")]

distance2GPO_by_DZN_CODE16 <-
  DZN_2016_centroids_tidy %>%
  .[id_by_city, on = "id"] %>%
  .[, distance2GPO := haversine_distance(lat, lon, GPO_lat, GPO_lon)] %>%
  .[, .(distance2GPO = min(distance2GPO),
        City = City[which.min(distance2GPO)]), keyby = c("DZN_CODE16")]

persons_by_GPO_distance_2011 <-
  distance2GPO_by_DZN_CODE11[DZN_2011_persons, on = "DZN_CODE11", nomatch=0L] %>%
  .[, Year := "2011"]

persons_by_GPO_distance_2016 <-
  distance2GPO_by_DZN_CODE16[DZN_2016_persons, on = "DZN_CODE16", nomatch=0L] %>%
  .[, Year := "2016"]

workers_by_GPO <-
  rbind(persons_by_GPO_distance_2011,
        persons_by_GPO_distance_2016,
        use.names = TRUE, fill = TRUE)

correspondence <-
  readxl::read_excel("../data-raw/cg_dzn_2011_dzn_2016.xls",
                     sheet = "Table 3",
                     range = "A6:D9562") %>%
  as.data.table %>%
  .[complete.cases(.)]
```


```{r workers_by_SLA2006_Home_Dest}
workers_by_SLA2006_Home_Dest <-
  fread("~/ABS-data/inbox/SLA-v-JTW-SLA-2006-clean.csv", header = FALSE, na.strings = "") %>%
  setnames(1:3, c("SLA_2006_Home", "SLA_2006_Work", "workers")) %>%
  drop_empty_cols %>%
  .[, SLA_2006_Home := zoo::na.locf(SLA_2006_Home, na.rm = FALSE)] %>%
  .[SLA_2006_Home != "Total"] %>%
  .[SLA_2006_Work != "Total"] %>%
  .[workers > 0]
```

```{r workers_by_SLA2006}
workers_by_SLA2006 <-
  workers_by_SLA2006_Home_Dest[, .(workers = sum(workers)), keyby = "SLA_2006_Work"]
```

```{r workersInnerSydney2006}

```

```{r}
SydneyMelJobGrowth <-
  read.table(text = " 	Sydney	Sydney 2	Blank	Melbourne	Melbourne 2
Inner 01	-2500	54500		23500	68000
Middle 01	13000	20000		26000	24500
Outer 01	35500	31000		56500	50000",
sep = "\t", header = TRUE, as.is = TRUE) %>%
  as.data.table %>%
  drop_empty_cols %>%
  setnames(c("Sydney", "Sydney.2"),
           c("Sydney_0106", "Sydney_0611")) %>%
  setnames(c("Melbourne", "Melbourne.2"),
           c("Melbourne_0106", "Melbourne_0611")) %>%
  setnames(1, "Ring") %>%
  melt.data.table(id.vars = "Ring", value.name = "delta_workers") %>%
  merge(data.table(Ring = c("Inner 01",
                            "Middle 01",
                            "Outer 01"),
                   distance_group = cut(c(1, 5, 25),
                                       breaks = c(-Inf, 2, 15, Inf), 
                                       labels = c("0-2km", "2-15km", ">15km"),
                                       ordered = TRUE)), 
        by = "Ring") %>%
  .[, City := "SYD"] %>%
  .[, Period := if_else(grepl("0611", variable), "2006-2011", "2001-2006")] %>%
  .[variable %pin% "Melbourne", City := "MEL"] %>%
  .[]
```

```{r workers_by_km_group_distance}
workers_by_km_group_distance <- 
  workers_by_GPO[distance2GPO < 35,
                 distance_group := cut(distance2GPO,
                                       breaks = c(-Inf, 2, 15, Inf), 
                                       labels = c("0-2km", "2-15km", ">15km"),
                                       ordered = TRUE)] %>%
  .[, .(total_workers = sum(persons)), keyby = c("Year", "City", "distance_group")] 
  
workers_by_km_group_distance %>%
  .[, .(delta_workers = last(total_workers) - first(total_workers)), keyby = c("City", "distance_group")] %>%
  .[, Period := "2011-2016"] %>%
  .[] %>%
  .[City %in% c("SYD", "MEL")] %>%
  rbind(SydneyMelJobGrowth, fill = TRUE, use.names = TRUE) %>%
  .[order(-Period)] %>%
  .[City %in% c("SYD", "MEL")] %>%
  .[order(City, distance_group, Period)] %>%
  grplot(aes(x = Period, y = delta_workers)) + 
  geom_bar(stat = "identity") +
  facet_grid(City ~ distance_group)
```

```{r}
workers_by_GPO %>% 
  .[distance2GPO < 2] %>% 
  .[City == "SYD"] %>% 
  .[DZN_2011_centroids_tidy, on = "DZN_CODE11", nomatch=0L] %>%
  ggplot(aes(lon, lat, size = persons)) +
  geom_point() +
  annotate_coastline("Sydney") +
  coord_map(xlim = c(151.19, 151.23), ylim = c(-33.89, -33.85)) +
  facet_wrap(~Year)
```


## Where have jobs grown (2011-2016)?


### Absolute change 

```{r n_employed_persons_2016_SYD_CBD}
n_employed_persons_2016_SYD_CBD <-
  JTW_2016_SA2 %>%
  .[D_SA2_NAME16 %ein% "Sydney - Haymarket - The Rocks"] %$%
  sum(workers) %>%
  round(-3)
```

```{r n_employed_persons_2011_SYD_CBD}
n_employed_persons_2011_SYD_CBD <-
  JTW_2011_SA2 %>%
  .[D_SA2_NAME11 %ein% "Sydney - Haymarket - The Rocks"] %$%
  sum(workers) %>%
  round(-1)
```


```{r n_employed_persons_2006_SYD_CBD}
n_employed_persons_2006_SYD_CBD <-
  JTW_2006_SLA %>%
  .[D_SLA_NAME06 %ein% "Sydney (C) - Inner"] %$%
  sum(workers)
```

```{r n_employed_persons_2001_SYD_CBD}
n_employed_persons_2001_SYD_CBD <- 186798
```

```{r n_employed_persons_1996_SYD_CBD}
n_employed_persons_1996_SYD_CBD <- 162609
```

```{r n_employed_persons_1991_SYD_CBD}
n_employed_persons_1991_SYD_CBD <-
  JTW_1991_SYD[Work_SLA %ein% "Sydney (Inner)"] %$% 
  sum(workers) %>%
  round(-3)
```


```{r ERP_by_ASGS_1991_2016}
ERP_by_ASGS_1991_2016 <-
  if (!file.exists("~/ABS-data/Estimated-Resident-Population-by-SA2-and-above-1991-2016.csv")) {
    fread("https://github.com/HughParsonage/ABS-data/raw/82c5be16745077b8c5049ae2c098c695f1e5ee5c/Estimated-Resident-Population-by-SA2-and-above-1991-2016.csv")
  } else {
    fread("~/ABS-data/Estimated-Resident-Population-by-SA2-and-above-1991-2016.csv")
  }
```

```{r AREA_by_SA2_2011}
AREA_by_SA2_2011 <-
  SA2_2011@data %>%
  as.data.table %>%
  .[, .(SA2_NAME11, ALBERS_SQKM = ALBERS_SQM / 1e6)]
```


```{r percent0}
percent0 <- function(x, d = 1) {
  paste0(formatC(100*x, digits = d, flag = " ", format = "f"), "%")
}
```

```{r delta_residents1116_density}
delta_residents1116_density <-
  ERP_by_ASGS_1991_2016 %>%
  .[REGIONTYPE == "SA2"] %>%
  .[Year %in% c(2011, 2016),
    .(Year, Region = NAME_2011, ERP = EstimatedResidentPopulation)] %>%
  dcast.data.table(Region ~ Year, value.var = "ERP") %>%
  .[`2011` > 0, .(SA2_NAME11 = Region,
                  `2011`,
                  `2016`,
                  delta_residents = (`2016` / `2011`) ^ (1/4) - 1)] %>%
  .[, breaks := .bincode(delta_residents,
                         breaks = quantile(delta_residents, probs = c(0:8)/8),
                         # breaks = c(-Inf, 200 * (0:6), Inf),
                         include.lowest = TRUE)] %>%
  .[, fillColor := c("white", gpal(7, reverse = TRUE))[breaks]] %>%
  .[, labelText := percent0(delta_residents)] %>%
  .[, labelTitle := SA2_NAME11] %>%
  .[]
```

```{r delta_residents1116}
delta_residents1116 <-
  ERP_by_ASGS_1991_2016 %>%
  .[REGIONTYPE == "SA2"] %>%
  .[Year %in% c(2011, 2016),
    .(Year, Region = NAME_2011, ERP = EstimatedResidentPopulation)] %>%
  dcast.data.table(Region ~ Year, value.var = "ERP") %>%
  .[`2011` > 0, .(SA2_NAME11 = Region,
                  `2011`,
                  `2016`,
                  delta_residents = (`2016` - `2011`))] %>%
  .[, p90 := quantile(delta_residents, probs = 0.9, na.rm = TRUE)] %>%
  
  .[delta_residents <= 0, fillColor := "white"] %>%
  .[delta_residents >= p90[1], fillColor := "black"] %>%
  .[is.na(fillColor),
    fillColor := gpal(6, reverse = TRUE)[.bincode(delta_residents,
                                                  breaks = ((0:5)/5) * p90[1], 
                                                  include.lowest = TRUE)]] %>%
  # .[complete.cases(.)] %>%
  .[, labelTitle := SA2_NAME11] %>%
  .[, labelText := scales::comma(delta_residents)] %>%
  .[]
```

```{r delta_residents_density}
delta_residents_density <-
  ERP_by_ASGS_1991_2016 %>%
  .[REGIONTYPE == "SA2"] %>%
  .[Year %in% c(2011, 2016), 
    .(Year, Region = NAME_2011, ERP = EstimatedResidentPopulation)] %>%
  .[AREA_by_SA2_2011, on = "Region==SA2_NAME11"] %>%
  .[, density := ERP / ALBERS_SQKM] %>%
  dcast.data.table(Region ~ Year, value.var = "density") %>%
  .[`2011` > 0, .(SA2_NAME11 = Region,
                  `2011`,
                  `2016`,
                  delta_residents = (`2016` / `2011`)^(1/4) - 1)] %>%
  .[, breaks := .bincode(delta_residents,
                         breaks = quantile(delta_residents, probs = c(0:8)/8),
                         # breaks = c(-Inf, 200 * (0:6), Inf),
                         include.lowest = TRUE)] %>%
  .[, fillColor := c("white", gpal(7, reverse = TRUE))[breaks]] %>%
  .[, labelText := paste0(if_else(delta_residents > 0, "+", ""),
                         percent(delta_residents), "<br>",
                         comma(floor(`2011`)), " --> ", comma(floor(`2016`)))] %>%
  .[, labelTitle := paste(SA2_NAME11, " density")] %>%
  .[]
```

```{r workers_by_SA2_Year}
workers_by_SA2_Year <-
  # Use 2011 boundaries
  rbindlist(list("2016" = JTW_2016_SA2[, .(workers = sum(workers)), keyby = .(SA2_NAME11 = D_SA2_NAME16)],
                 "2011" = JTW_2011_SA2[, .(workers = sum(workers)), keyby = .(SA2_NAME11 = D_SA2_NAME11)]),
            idcol = "Year") %>%
  dcast.data.table(SA2_NAME11 ~ Year, value.var = "workers")
```

```{r workers_by_Sydney_SA3_Year}
workers_by_Sydney_SA3_Year <- 
  data.table(Year = c(1991, 2016),
             workers = c(359795, 494315)) %>%
  .[, AREA := 25.0641]
```

```{r delta_workers1116}
delta_workers1116 <- workers_by_SA2_Year[, .(SA2_NAME11, delta_workers = `2016` - `2011`)]
```

```{r delta_residents_vs_workers}
delta_residents_vs_workers <-
  delta_workers1116[delta_residents1116, on = "SA2_NAME11", nomatch=0L]
```

```{r delta_workers_0616_abs}
delta_workers_0616_abs <-
  JTW_2006_SLA
```


```{r SA2_2016_residents}
SA2__persons <- 
  SA2__Persons %>%
  SA2016_decoder[., on = "SA2_MAIN16==SA2_MAINCODE_2016"] %>%
  set_cols_first("SA2_NAME16") %>%
  .[, fillColor := gpal(7, reverse = TRUE)[dplyr::ntile(persons, 7)]] %>%
  .[, labelTitle := SA2_NAME16] %>%
  .[, labelText := scales::comma(persons)]
```

```{r n_workers_by_Year_SYD_CBD}
n_workers_by_Year_SYD_CBD <-
  data.table(Year = seq(from = 1991, to = 2016, by = 5), 
             workers = as.numeric(unlist(mget(x = paste0("n_employed_persons_", seq(1991, 2016, by = 5), "_SYD_CBD")))))
```

```{r SA3_2011_centroids}
SA3_2011_centroids <- 
  cbind(SA3_2011@data, gCentroid(SA3_2011, byid = TRUE)@coords) %>%
  as.data.table
```

```{r SA3NAME_vs_dist2SYDGPO}
SA3NAME_vs_dist2SYDGPO <-
  SA3_2011_centroids %>%
  .[, dist2SYDGPO := haversine_distance(y, x, SYD$LATITUDE, SYD$LONGITUDE)] %>%
  .[]
```

```{r SA2NAME_vs_dist2SYDGPO}
SA2NAME_vs_dist2SYDGPO <-
  SA2_2011_weighted_centroids %>%
  .[, SA2_NAME11 := latlon2SA(lat, lon, to = "SA2", yr = "2011", return = "v")] %>%
  .[, dist2SYDGPO := haversine_distance(lat, lon, SYD$LATITUDE, SYD$LONGITUDE)] %>%
  .[]
```


```{r pop-inner-ring-vs-CBD-employment}
ERP_by_ASGS_1991_2016 %>%
  .[REGIONTYPE == "SA3"] %>%
  .[NAME_2011 %in% SA3NAME_vs_dist2SYDGPO[dist2SYDGPO < 10][["SA3_NAME11"]]] %>%
  ggplot(aes(x = Year, y = EstimatedResidentPopulation, group = NAME_2011, color = NAME_2011)) + 
  geom_line()
```

```{r SydneyInnerCitySA3_minus_CBD_by_Year}
SydneyInnerCitySA3_minus_CBD_by_Year <-
  ERP_by_ASGS_1991_2016 %>%
  .[NAME_2011 %ein% c("Sydney Inner City", "Sydney - Haymarket - The Rocks")] %>%
  .[, .(NAME_2011, Year, EstimatedResidentPopulation)] %>%
  dcast.data.table(Year ~ NAME_2011, value.var = "EstimatedResidentPopulation") %>%
  .[, `Sydney Inner City minus CBD` := `Sydney Inner City` - `Sydney - Haymarket - The Rocks`] %>%
  .[, .(Year, `Sydney - Haymarket - The Rocks`, `Sydney Inner City minus CBD`)] %>%
  melt.data.table(id.vars = "Year",
                  variable.name = "NAME_2011",
                  variable.factor = FALSE,
                  value.name = "EstimatedResidentPopulation")
```

```{r pop-inner-ring-vs-CBD-employment-CBD-split-out}
ERP_by_ASGS_1991_2016 %>%
  .[REGIONTYPE == "SA3"] %>%
  .[NAME_2011 %in% SA3NAME_vs_dist2SYDGPO[dist2SYDGPO < 10][["SA3_NAME11"]]] %>%
  # .[NAME_2011 %enotin% "Sydney Inner City"] %>%
  # rbind(SydneyInnerCitySA3_minus_CBD_by_Year, use.names = TRUE, fill = TRUE) %>%
  # .[NAME_2011 %enotin% c("Sydney - Haymarket - The Rocks")] %>%
  .[, Index := EstimatedResidentPopulation / first(EstimatedResidentPopulation),
    keyby = "NAME_2011"] %>%
  .[, last_pop := last(Index), keyby = "NAME_2011"] %>%
  .[, SA3 := reorder(NAME_2011, -last_pop)] %>%
  ggplot(aes(x = Year, y = Index, group = SA3, color = SA3)) + 
  geom_line() + 
  scale_y_continuous(labels = function(x) 100 * x,
                     breaks = c(1, 1.5, 2, 2.5))
```


```{r pop_index-vs-Year-SA2-by-distance--Sydney}
# ERP_by_ASGS_1991_2016
```

```{r slope}
slope <- function(x, y) {
  x <- as.double(x)
  y <- as.double(y)
  n <- length(x)
  {sum(x * y) - sum(x) * sum(y) / n} /
    {sum(x^2) - sum(x)^2 / n}
}
```

```{r SA2011_decoder}
SA2011_decoder <- as.data.table(SA2_2011)
```

```{r growth_by_SA2}
ERP_by_ASGS_1991_2016 %>%
  .[NAME_2011 %ein% SA2011_decoder[GCC_NAME11 == "Greater Sydney"][["SA2_NAME11"]]] %>%
  .[SA2_2011_weighted_centroids, on = "CODE_2011==SA2_MAIN11", nomatch=0L] %>%
  .[, distance2GPO := haversine_distance(lat, lon, SYD$LATITUDE, SYD$LONGITUDE)] %>%
  .[] %>%
  .[, distance_group := cut(distance2GPO, breaks = c(-Inf, 2, 10, 20, Inf), 
                            labels = c("<2km", "2-10km", "10-20km", "20km+"),
                            include.lowest = TRUE,
                            ordered_result = TRUE)] %>%
  .[EstimatedResidentPopulation > 1000] %>%
  .[, Index := EstimatedResidentPopulation / first(EstimatedResidentPopulation),
    keyby = "NAME_2011"] %>%
  # Some SA2s have zero population.
  .[!is.nan(Index)] %>%
  .[, .(Year, Index, NAME_2011, distance_group)] %>%
  grplot(aes(x = Year, y = Index, group = NAME_2011, color = distance_group)) + 
  geom_line() +
  geom_smooth(aes(group = distance_group)) +
  facet_wrap(~distance_group)
```

```{r pal_v}
#' @return Viridis palette of x by quantile (decile by default).
palette_v <- function(x, n = 10) {
  input <- data.table(x = x)
  input[, order := 1:.N]
  input[, decile := ntile(x, n)]
  
  colortbl <- data.table(decile = 1:10,
                         colour = viridis(10))
  
  colortbl[input, on = "decile"] %>%
    setorder(order) %>%
    .[["colour"]]
}

pal_v <- colorNumeric(palette = viridis(10),
                      domain = NULL)
```

```{r gpal_v}
gpal_v <- colorNumeric(palette = c("white", gpal(7, reverse = TRUE)),
                      domain = NULL)
```

```{r slope_by_SA2}
slope_by_SA2 <- 
  ERP_by_ASGS_1991_2016 %>%
  .[Year > 2005] %>%
  .[NAME_2011 %ein% SA2011_decoder[GCC_NAME11 == "Greater Sydney"][["SA2_NAME11"]]] %>%
  .[SA2_2011_weighted_centroids, on = "CODE_2011==SA2_MAIN11", nomatch=0L] %>%
  .[, distance2GPO := haversine_distance(lat, lon, SYD$LATITUDE, SYD$LONGITUDE)] %>%
  .[, .(Slope = slope(Year, EstimatedResidentPopulation)), keyby = "NAME_2011"] %>%
  .[, fillColor := gpal_v(rank(Slope))] %>%
  .[, labelTitle := NAME_2011] %>%
  .[, labelText := scales::comma(floor(Slope))] %>%
  .[, SA2_NAME11 := NAME_2011] %>%
  set_cols_first("SA2_NAME11") %>%
  .[]
```

```{r slope-by-SA2}
grattan_leaflet(slope_by_SA2, "2011", simple = TRUE)
```

```{r pop_density_by_SA2_by_Year}
density_by_SA2_by_Year <-
  as.data.table(SA2_2011) %>%
  .[ERP_by_ASGS_1991_2016, on = "SA2_NAME11==NAME_2011", nomatch=0L] %>%
  .[SA2_NAME11 %ein% SA2011_decoder[GCC_NAME11 == "Greater Sydney"][["SA2_NAME11"]]] %>%
  .[, pop_density := EstimatedResidentPopulation / (ALBERS_SQM / 1e6)] %>%
  .[]
```

```{r pop_density_by_SA3_by_Year}
density_by_SA3_by_Year <-
  as.data.table(SA3_2011) %>%
  .[ERP_by_ASGS_1991_2016, on = "SA3_NAME11==NAME_2011", nomatch=0L] %>%
  .[SA3_NAME11 %ein% SA2011_decoder[GCC_NAME11 == "Greater Sydney"][["SA3_NAME11"]]] %>%
  .[, pop_density := EstimatedResidentPopulation / (ALBERS_SQM / 1e6)] %>%
  .[]
```

```{r slopeDensity_by_SA2_1991_2016}
slopeDensity_by_SA2_1991_2016 <-
  density_by_SA2_by_Year %>%
  .[, .(Slope_density = slope(Year, pop_density)), keyby = "SA2_NAME11"] %>%
  .[, Slope_density := round(Slope_density)] %>%
  .[, fillColor := gpal_v(rank(Slope_density))] %>%
  .[Slope_density <= 0, fillColor := "white"] %>%
  .[, labelTitle := SA2_NAME11] %>%
  .[, labelText := paste(scales::comma(floor(Slope_density)),
                         "persons per sqkm per year")] %>%
  .[]
  

```

```{r density-increase-1991-2016}
grattan_leaflet(slopeDensity_by_SA2_1991_2016, Year = "2016", simple = TRUE)
```


```{r slopeDensity_by_SA3_1991_2016}
slopeDensity_by_SA3_1991_2016 <-
  density_by_SA3_by_Year %>%
  .[, .(Slope_density = slope(Year, pop_density)), keyby = "SA3_NAME11"] %>%
  .[, Slope_density := round(Slope_density)] %>%
  .[, fillColor := gpal_v(Slope_density)] %>%
  .[Slope_density <= 0, fillColor := "white"] %>%
  .[, labelTitle := SA3_NAME11] %>%
  .[, labelText := paste(scales::comma(floor(Slope_density)),
                         "persons per sqkm per year")] %>%
  .[]
```

```{r density-increase-1991-2016-SA3, fig.cap="Outside the centre of Sydney, density has not grown much"}
grattan_leaflet(slopeDensity_by_SA3_1991_2016, Year = "2011", simple = TRUE)
```

```{r SALM}
SALM <- 
  if (file.exists("~/data-employment-gov-au/salm_smoothed_sa2_datafile_june_quarter_2017.csv")) {
    fread("~/data-employment-gov-au/salm_smoothed_sa2_datafile_june_quarter_2017.csv")
  } else {
    tryCatch({
    fread("https://docs.employment.gov.au/system/files/doc/other/salm_smoothed_sa2_datafiles_-_june_quarter_2017.csv") %>%
      melt.data.table(id.vars = c("Data item",
                                  "Statistical Area Level 2 (SA2)",
                                  "SA2 Code"),
                      variable.factor = FALSE,
                      variable.name = "Month") %>%
      .[, Date := as.Date(paste0("01-", Month), format = "%d-%b-%y")] %>%
      setnames(c("Statistical Area Level 2 (SA2)", "SA2 Code"),
               c("SA2_NAME16", "SA2_MAIN16")) %>%
      .[, .(SA2_NAME16, SA2_MAIN16, Date, Item = `Data item`, value)] %T>%
      fwrite("~/data-employment-gov-au/salm_smoothed_sa2_datafile_june_quarter_2017.csv") %>% 
      .[]
    }, error = function(e) {
    fread("https://github.com/HughParsonage/data-employment-gov-au/raw/101c7dc2f137e31b8f4de5669f8653340784c0f9/salm_smoothed_sa2_datafile_june_quarter_2017.csv")
    })
  }
```


```{r slopeDensity_by_SA2_2006_2016}
slopeDensity_by_SA2_2006_2016 <-
  density_by_SA2_by_Year %>%
  .[Year >= 2006] %>%
  .[, .(Slope_density = slope(Year, pop_density)), keyby = "SA2_NAME11"] %>%
  .[, Slope_density := round(Slope_density)] %>%
  .[, fillColor := gpal_v(rank(Slope_density))] %>%
  .[Slope_density <= 0, fillColor := "white"] %>%
  .[, labelTitle := SA2_NAME11] %>%
  .[, labelText := paste(scales::comma(floor(Slope_density)), "persons per sqkm per year")] %>%
  .[]
  

```




Australia's economic future lies in its cities.
But city planning is moving jobs and residential areas in different directions.
The city sprawl is a well-understood phenomenon but what is less well-known is that jobs are becoming more concentrated.
The modern economy more than ever thrives with a high concentration of people working in the same square kilometre, 
but the modern planning system limits how close they can live to these hot spots.

In 1991, `r comma(n_employed_persons_1991_SYD_CBD)` worked in the Sydney CBD.
Twenty years later, `r comma(n_employed_persons_2011_SYD_CBD)` did and in the most recent census, `r comma(n_employed_persons_2016_SYD_CBD)` of the population's main job was located in the 
`r SA2_2016@data %>% as.data.table %>% .[SA2_NAME16 %pin% "The Rocks$"] %>% .[["AREASQKM16"]] %>% round(1)` km^2.
at the heart of the city.

```{r SCG_area}
SCG_area <- 0.0186 # sqkm
WatsonsBaySlope <-
  slopeDensity_by_SA2_1991_2016 %>%
  .[SA2_NAME11 %ein% "Rose Bay - Vaucluse - Watsons Bay"] %>%
  .[["Slope_density"]]
```

Over this period, with the exception of the Sydney CBD proper, the population of the inner ring grew modestly.
The density of the Eastern Suburbs from Rose Bay to Watsons Bay increased by a mere 13 persons per square kilometre per year, the same rate of density increase as one person walking on to the SCG every `r floor(1/(WatsonsBaySlope * SCG_area))` years.

By contrast the density of jobs in the Sydney Inner City SA3 increased by `r round(workers_by_Sydney_SA3_Year %$% {(last(workers) - first(workers))  / AREA[1] / 25})` workers per square kilometre per year and the Sydney CBD proper by 1600 per square kilometre per year, or 181,000 in 25 years. 

```{r n_jobs_SYD_2016}
n_jobs_SYD_2016 <-
  JTW_2016_SA2 %>%
  .[SA2016_decoder, on = "D_SA2_NAME16==SA2_NAME16"] %>%
  .[, .(workers = sum(workers, na.rm = TRUE)), keyby = "GCC_NAME16"] %>%
  .[GCC_NAME16 == "Greater Sydney"] %>%
  .[["workers"]]
```

The Greater Metropolitan region of Sydney had 1.38~million jobs.
This increased to `r comma(n_jobs_SYD_2016)`. 






